Manifold Clustering for Latent Structure Block Models

### 2022-02-23 draft/draft.Rmd
update to draft

```{r}
import::from(magrittr, `%>%`, `%<>%`)
library(ggplot2)
source('../functions.R')
source('~/dev/pabm-grdpg/functions.R')
source('https://mtrosset.pages.iu.edu/Courses/675/graph.r')

@@

doMC::registerDoMC(8)
```

@@

```{r, cache = TRUE}
random.starts <- 4
eps <- 1e-3
out <- lapply(seq(random.starts), function(i) {
  set.seed(i)
  intercept <- (i / random.starts > .5 + eps)
  manifold.clustering(Xhat, 2, parallel = TRUE, intercept = intercept, 
                      maxit = 50,
                      verbose = FALSE)
})
names(out) <- paste('random', seq(random.starts))
names(out) <- sapply(seq(random.starts), function(i) {
  if (i / random.starts < .5 + eps) {
    paste('random', i, '\n(w/o intercept)')
  } else {
    paste('random', i, '\n(w/ intercept)')
  }
})

# out[['random\n(w/ intercept)']] <- manifold.clustering(Xhat, 2,
#                                           parallel = TRUE,
#                                           intercept = TRUE, 
#                                           maxit = 50, 
#                                           initialization = 'random',
#                                           verbose = FALSE)

out[['spectral clust\n(w/o intercept)']] <- manifold.clustering(
  Xhat, 2, 
  parallel = TRUE, 
  intercept = FALSE, 
  maxit = 50, 
  initialization = 'spectral',
  verbose = FALSE)

# out[['ground truth']] <- manifold.clustering(Xhat, 2,
#                                             parallel = TRUE,
#                                             intercept = FALSE,
#                                             maxit = 50,
#                                             initialization = z,
#                                             verbose = FALSE)

out[['spectral clust\n(w/ intercept)']] <- manifold.clustering(
  Xhat, 2,
  parallel = TRUE,
  intercept = TRUE,
  maxit = 50,
  initialization = 'spectral',
  verbose = FALSE)
```

@@

loss.df <- plyr::ldply(seq_along(out), function(i) {
  dplyr::data_frame(loss = out[[i]]$loss,
                    initialization = factor(names(out),
                                            levels = names(out))[i]) %>%
    dplyr::mutate(iteration = seq(dplyr::n()))
})

ggplot(loss.df) + 
  geom_point(aes(x = iteration, y = loss, colour = initialization)) + 
  geom_line(aes(x = iteration, y = loss, colour = initialization)) + 
  # ylim(0, max.loss) + 
  labs(y = 'clustering loss')
```

@@

>>>>

```{r}
import::from(magrittr, `%>%`, `%<>%`)
library(ggplot2)
library(gganimate)
source('../functions.R')
source('~/dev/pabm-grdpg/functions.R')
source('https://mtrosset.pages.iu.edu/Courses/675/graph.r')

@@

doMC::registerDoMC(parallel::detectCores())
```

@@

```{r, cache = TRUE}
maxit <- 50
animate <- TRUE
animation.dir <- '~/dev/manifold-block-models/draft'
random.starts <- 2
eps <- 1e-3
out <- lapply(seq(random.starts), function(i) {
  set.seed(i)
  intercept <- (i / random.starts > .5 + eps)
  manifold.clustering(Xhat, 2, 
                      parallel = TRUE, 
                      intercept = intercept, 
                      maxit = maxit,
                      eps = 1e-3,
                      verbose = FALSE,
                      animate = animate,
                      animation.dir = animation.dir,
                      animation.title = i)
})
names(out) <- paste('random', seq(random.starts))
names(out) <- sapply(seq(random.starts), function(i) {
  if (i / random.starts < .5 + eps) {
    paste('random', i, '\n(w/o intercept)')
  } else {
    paste('random', i, '\n(w/ intercept)')
  }
})

out[['spectral clust\n(w/o intercept)']] <- manifold.clustering(
  Xhat, 2, 
  parallel = TRUE, 
  intercept = FALSE, 
  maxit = maxit, 
  eps = 1e-3,
  initialization = 'spectral',
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'spec-clust-no-intercept')

out[['spectral clust\n(w/ intercept)']] <- manifold.clustering(
  Xhat, 2,
  parallel = TRUE,
  intercept = TRUE,
  maxit = maxit,
  eps = 1e-3,
  initialization = 'spectral',
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'spec-clust-w-intercept')

out[['ground truth\n(w/o intercept)']] <- manifold.clustering(
  Xhat, 2,
  parallel = TRUE,
  intercept = FALSE,
  maxit = maxit,
  eps = 1e-3,
  initialization = z,
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'groundtruth-no-intercept')

out[['ground truth\n(w/ intercept)']] <- manifold.clustering(
  Xhat, 2,
  parallel = TRUE,
  intercept = TRUE,
  maxit = maxit,
  eps = 1e-3,
  initialization = z,
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'groundtruth-w-intercept')
```

@@

loss.df <- plyr::ldply(seq_along(out), function(i) {
  dplyr::tibble(loss = out[[i]]$loss,
                initialization = factor(names(out),
                                        levels = names(out))[i]) %>%
    dplyr::mutate(iteration = seq(dplyr::n()))
})

ggplot(loss.df) + 
  geom_point(aes(x = iteration, y = loss, colour = initialization)) + 
  geom_line(aes(x = iteration, y = loss, colour = initialization)) + 
  # ylim(0, max.loss) + 
  # scale_y_log10() + 
  labs(y = 'clustering loss')
```

@@

\begin{example}

# Simulation Study
